---
title: "Multi Level and Quantile Models in BRMS"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

A notebook which is testing the implementation of 
BRMS quantile hierarchical models using generated
datasets.


### No Relationship

First lets generate some data where there is absolutely
no relationship to it at all. We will fit a normal
linear model. 

$$
y_i\;{\sim}\;\mathcal{N}(\mu_i, \sigma_i)
$$
In this instance, we will assume no relationshi

We will then fit some quantile models. In the quantile bayesian
quantile regression, we fit an asymmetric Laplace distribution:

$$
y_i\;=\;Laplace(m_i, \lambda_i,p_i)
$$
$m$ is a location parameter, $\lamda$ is a scale parameter, and $p$ is a quantile/percentile
parameter, representing the quantile of interest. When used for Bayesian inference, the
asymmetric laplace funtion displays the same properties as the loss function used in quantile
regression. 
$$
\displaystyle f(x;m,\lambda ,p)={\frac {p(1-p)}{\lambda }}{\begin{cases}\exp \left(-((p-1)/\lambda )(x-m)\right)&{\text{if }}x\leq m\\[4pt]\exp(-(p/\lambda )(x-m))&{\text{if }}x>m\end{cases}}
$$

We will visualise the results and see how they compare.

```{r, message=F, warning=FALSE}

library(ggplot2)
library(ggExtra)
library(brms)
library(bayesplot)
library(tidybayes)
library(magrittr)
library(tidyr)
library(dplyr)

number_of_samples <- 500

# Generating random points
# Using skew normal distribution, but skewness is 0
X <- brms::rskew_normal(n=number_of_samples,mu = 10,sigma = 1,alpha = 0)
y <- brms::rskew_normal(n=number_of_samples,mu = 2,sigma = 2,alpha = 0)

df <- tibble::as_tibble(list(
  x=X,
  y=y
))



plot <- ggplot(df, aes(x,y))+
  geom_point()

plot <- ggMarginal(plot,type="histogram")
plot

```
##

```{r, results='hide'}
m <- brms::brm(y~x,data = df)



```
```{r}
fitted_values <- fitted(m,dpar="mu") %>% as_tibble()

fitted_values$x <- df$x
ggplot(df, aes(x,y))+
  geom_point()+
  geom_ribbon(data=fitted_values,mapping = aes(x=x, ymin=Q2.5 ,ymax=Q97.5),fill="dodgerblue4", alpha=0.3)+
  geom_line(data=fitted_values,mapping = aes(x=x, y=Estimate))

```

```{r results='hide'}
m_q0.25 <- brms::brm(bf(y~x, 
                        quantile=0.25),
                     data = df,
                     
                     family=asym_laplace())

m_q0.75 <- brms::brm(bf(y~x, 
                        quantile=0.75),
                     data=df,
                     family=asym_laplace())
```


```{r}

fitted_values_q0.25 <- fitted(m_q0.25,dpar="mu") %>% as_tibble()
fitted_values_q0.25$x <- df$x
fitted_values_q0.75 <- fitted(m_q0.75,dpar="mu") %>% as_tibble()
fitted_values_q0.75$x <- df$x

fitted_values$x <- df$x

ggplot(df, aes(x,y))+
  geom_point()+
  
  geom_ribbon(data=fitted_values,
              mapping = aes(x=x, ymin=Q2.5 ,ymax=Q97.5, fill="Mean"),
              
              alpha=0.3)+
  
  geom_line(data=fitted_values,mapping = aes(x=x, y=Estimate))+
  
  geom_ribbon(data=fitted_values_q0.25,
              mapping = aes(x=x, ymin=Q2.5 ,ymax=Q97.5, fill="Q0.25"),
              
              alpha=0.3)+
  
  geom_line(data=fitted_values_q0.25,mapping = aes(x=x, y=Estimate))+
  
  geom_ribbon(data=fitted_values_q0.75,
              mapping = aes(x=x, ymin=Q2.5 ,ymax=Q97.5,  fill="Q0.75"),
              alpha=0.3)+
  
  
  geom_line(data=fitted_values_q0.75,mapping = aes(x=x, y=Estimate))+
  scale_fill_hue(labels = c("Mean", "Q0.25", "Q0.75"))






```

```{r}

number_of_samples <- 500

X <- cbind(1, runif(number_of_samples, 0, 100)) # Generate N uniform random numbers between 0-100

B_mu <- c(10, 0.2)
B_sigma <- c(0, 0.022)
B_alpha <- c(0, 0)


mu <- X %*% B_mu # Matrix Multiplication
sigma <- exp(X %*% B_sigma)
alpha <- exp(X %*% B_alpha)


y <- rnorm(N, mu, sigma)
sim_data <- data.frame(y, X)

# Generating random points
# Using skew normal distribution, but skewness is 0
y <- brms::rskew_normal(n=number_of_samples,mu = mu,sigma = sigma,alpha = alpha)

df <- tibble::as_tibble(list(
  x=X[,2],
  y=y
))



plot <- ggplot(df, aes(x,y))+
  geom_point()

plot <- ggMarginal(plot,type="histogram")
plot



```
```{r results='hide'}
m <- brms::brm(y~x,data = df)
m_q0.25 <- brms::brm(bf(y~x, 
                        quantile=0.25),
                     data = df,
                     
                     family=asym_laplace())

m_q0.75 <- brms::brm(bf(y~x, 
                        quantile=0.75),
                     data=df,
                     family=asym_laplace())

```


```{r}

fitted_values_m <- fitted(m,dpar="mu") %>% as_tibble()
fitted_values_m$x <- df$x
fitted_values_q0.25 <- fitted(m_q0.25,dpar="mu") %>% as_tibble()
fitted_values_q0.25$x <- df$x
fitted_values_q0.75 <- fitted(m_q0.75,dpar="mu") %>% as_tibble()
fitted_values_q0.75$x <- df$x


ggplot(df, aes(x,y))+
  geom_point()+
  
  geom_ribbon(data=fitted_values_m,
              mapping = aes(x=x, ymin=Q2.5 ,ymax=Q97.5, fill="Mean"),
              
              alpha=0.3)+
  
  geom_line(data=fitted_values_m,mapping = aes(x=x, y=Estimate))+
  
  geom_ribbon(data=fitted_values_q0.25,
              mapping = aes(x=x, ymin=Q2.5 ,ymax=Q97.5, fill="Q0.25"),
              
              alpha=0.3)+
  
  geom_line(data=fitted_values_q0.25,mapping = aes(x=x, y=Estimate))+
  
  geom_ribbon(data=fitted_values_q0.75,
              mapping = aes(x=x, ymin=Q2.5 ,ymax=Q97.5,  fill="Q0.75"),
              alpha=0.3)+
  
  
  geom_line(data=fitted_values_q0.75,mapping = aes(x=x, y=Estimate))+
  scale_fill_hue(labels = c("Mean", "Q0.25", "Q0.75"))



```
# Generating Multi Level Data

```{r}



#' Generate Multi-Level Data
#'
#' @param n 
#' @param intercept 
#' @param level_1_intercept 
#' @param level_1_sd 
#' @param level_2_intercept 
#' @param level_2_sd 
#' @param sigma 
#'
#' @return
#' @export
#'
#' @examples
gen_lg_data <- function(
  n,
  intercept,
  
  level_1_intercept,
  level_1_sd,
  
  level_2_intercept,
  level_2_sd,
  
  sigma,
    J, cs, gam, G, sigma2 = 1
    
    ) {
  
  
  
  X <- cbind(1, seq_len(cs) - 1)  # for each individual
  X <- X[rep(seq_len(cs), J), ]  # repeat each row cs times
  pid <- seq_len(J)  # individual id
  pid <- rep(pid, each = cs)  # repeat each ID cs times
  # Generate person-level (lv-2) random effects
  uj <- rmnorm(J, mean = rep(0, 2), varcov = G)
  # Generate repeated-measure-level (lv-1) error term
  eij <- rnorm(J * cs, sd = sqrt(sigma2))
  # Compute beta_j's
  betaj <- matrix(gam, nrow = J, ncol = 2, byrow = TRUE) + uj
  # Compute outcome:
  y <- rowSums(X * betaj[pid, ]) + eij
  # Form a data frame
  sim_dat1 <- tibble(y, time = X[ , 2], pid)
  # Return data
  return(sim_dat1)
}

number_of_samples <- 500

X <- cbind(1, runif(number_of_samples, 0, 100)) # Generate N uniform random numbers between 0-100

B_mu <- c(10, 0.2)
B_sigma <- c(0, 0.022)
B_alpha <- c(0, 0)


mu <- X %*% B_mu # Matrix Multiplication
sigma <- exp(X %*% B_sigma)
alpha <- exp(X %*% B_alpha)


y <- rnorm(N, mu, sigma)
sim_data <- data.frame(y, X)

# Generating random points
# Using skew normal distribution, but skewness is 0
y <- brms::rskew_normal(n=number_of_samples,mu = mu,sigma = sigma,alpha = alpha)

df <- tibble::as_tibble(list(
  x=X[,2],
  y=y
))



plot <- ggplot(df, aes(x,y))+
  geom_point()

plot <- ggMarginal(plot,type="histogram")
plot



```